---
title: "measure variable responces of categorical feature with DALEX + mlr"
author: "Satoshi Kato"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  html_document:
    fig_caption: yes
    pandoc_args:
    - --from
    - markdown+autolink_bare_uris+tex_math_single_backslash-implicit_figures
    toc: yes
    toc_depth: 4
    keep_md: yes
  word_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
require(tidyverse)
require(mlr)
require(iml)

knitr::opts_knit$set(progress = TRUE, 
                     verbose = TRUE, 
                     root.dir = ".")

knitr::opts_chunk$set(collapse = FALSE, 
                      prompt  = FALSE,
                      comment = "", 
                      message = TRUE, 
                      warning = FALSE, 
                      echo=TRUE)
set.seed(12345)
```

# read mlr models

regression task for apartments dataset.

```{r mlr.prep, message=FALSE}
tuned.model <- readRDS("./tuned_models.RDS")
# tuned.model %>% str(2)

```

# DALEX + mlr

according to:

https://rawgit.com/pbiecek/DALEX_docs/master/vignettes/DALEX_mlr.html

# The explain() function

## custom predict()

For the models created by mlr package we have to provide custom predict function which takes two arguments: model and newdata and returns a numeric vector with predictions because function predict() from mlr returns not only predictions but an object with more information.

```{r}
predictMLR <- function(object, newdata) {
  pred <- predict(object, newdata=newdata)
  response <- pred$data$response
  return(response)
}

```

## simple plot

```{r}
require(breakDown)
data("apartmentsTest", package = "DALEX")
X <- apartmentsTest[,-1]

target <- 1

br.down.rf <- broken(
  model = tuned.model[["rf"]], 
  data  = X,
  new_observation = X[target, ], 
  direction = "down",
  predict.function = predictMLR,
  keep_distributions=TRUE)

pbr.down.rf <- plot(br.down.rf) + ggtitle("direction = DOWN")

```
## plot  distributions of conditional propotions

* 	if `keep_distributions = TRUE`, then the distribution of partial predictions is stored in addition to the average.
* 	if `plot_distributions = TRUE`, if TRUE then distributions of conditional propotions will be plotted. This requires keep_distributions=TRUE in the broken.default().


```{r}
plot(br.down.rf, plot_distributions =TRUE)

```

## up and down

```{r, fig.height=6, fig.width=9}
br.up.rf <- broken(
  model = tuned.model[["rf"]], 
  data  = X,
  new_observation = X[target, ], 
  direction = "up",
  predict.function = predictMLR,
  keep_distributions=TRUE)

pbr.up.rf <- plot(br.down.rf) + ggtitle("direction = UP")

gridExtra::grid.arrange(pbr.down.rf, pbr.up.rf, ncol=2)

```


## multiple explainers

```{r, fig.height=9, fig.width=6}
pbr.down <- list()

for(model.name in names(tuned.model)){
  
  br.down <- broken(
    model = tuned.model[[model.name]], 
    data  = X,
    new_observation = X[target, ], 
    direction = "down",
    predict.function = predictMLR,
    keep_distributions=TRUE)
  
  pbr.down[[model.name]] <- plot(br.down) + ggtitle(model.name)
}

gridExtra::grid.arrange(grobs = pbr.down, ncol=1)


```

